---
title: "water_balance_greenhouse"
author: "iael"
date: "2024-01-07"
output: html_document
---

```{r setup, include=FALSE}
library("grid")
library("tidyverse")
library("lubridate")
library("gridExtra")
library("cowplot")
library("readxl")
library("ggpubr")
library("rstatix")
library("tidyr")
library("multcompView")
library("scales")
library("dplyr")
library("ggpattern")
library("RColorBrewer")

```



```{r include=FALSE}
raw_data <- read_excel("onion_analysis.xlsx", sheet = "data")
dates<-colnames(raw_data)
dates<-dates[6:length(dates)]
date_object <- as.Date(as.numeric(dates), origin = "1899-12-30")

irr_drain<-pivot_longer(raw_data, cols = 6:ncol(raw_data), names_to ="dateN",
 values_to = "mlday")
irr_drain$time<-as.Date(as.numeric(irr_drain$dateN), origin = "1899-12-30")
irr_drain$TREATMENT<-as.factor(irr_drain$TREATMENT)
irr_drain$Date <- as.POSIXct(irr_drain$time)
irr_drain$N_level_p<-irr_drain$N_level*100


raw_data_irr <- read_excel("onion_analysis.xlsx", sheet = "irrECN")
raw_data_irr$Date <- as.POSIXct(raw_data_irr$date)
raw_data_irr$treatment<-as.factor(raw_data_irr$treatment)
raw_data_irr$N_level_p<-as.factor(as.numeric(raw_data_irr$Nlevel)*100)
raw_data_irr$Nlevel<-as.factor(raw_data_irr$Nlevel)
raw_data_irr$targetEC.n<-as.factor(raw_data_irr$targetEC)

raw_data_drainN <- read_excel("onion_analysis.xlsx", sheet = "drainN")
dates<-colnames(raw_data_drainN)
dates<-dates[6:length(dates)] #not sure if 5 or 6
date_object <- as.Date(as.numeric(dates), origin = "1899-12-30")
#### here working on this data!!!!!
drainN<-pivot_longer(raw_data_drainN, cols = 6:ncol(raw_data_drainN), names_to ="dateN",
 values_to = "drainN_NO3mgl")
drainN$time<-as.Date(as.numeric(drainN$dateN), origin = "1899-12-30")
drainN$TREATMENT<-as.factor(drainN$TREATMENT)
drainN$Date <- as.POSIXct(drainN$time)
drainN$N_level_p<-drainN$N_level*100

raw_data_drainEC <- read_excel("onion_analysis.xlsx", sheet = "drainEC")
dates<-colnames(raw_data_drainEC)
dates<-dates[6:length(dates)] #not sure if 5 or 6
date_object <- as.Date(as.numeric(dates), origin = "1899-12-30")
#### here working on this data!!!!!
drainEC<-pivot_longer(raw_data_drainEC, cols = 6:ncol(raw_data_drainEC), names_to ="dateN",
 values_to = "drainEC")
drainEC$time<-as.Date(as.numeric(drainEC$dateN), origin = "1899-12-30")
drainEC$Date <- as.POSIXct(drainEC$time)

drainEC$TREATMENT<-as.factor(drainEC$TREATMENT)
drainEC$N_level_p<-drainEC$N_level*100


#join drainEC and drainN so that we can analyze them together 
drain_NEC<-full_join(drainN,drainEC,by=c("Date","TREATMENT","EC","N_level","Bucket #","code","dateN","time","N_level_p"))


drain_NEC_avg <- drain_NEC %>%                         # Aggregate data
  group_by(Date, TREATMENT) %>% 
  dplyr::summarize(drainN_NO3mgl = mean(drainN_NO3mgl),drainEC=mean(drainEC),N_level_p=mean(N_level_p),EC=mean(EC)) %>% 
  as.data.frame()
drain_NEC_avg<-na.omit(drain_NEC_avg)



u_date<-unique(irr_drain$time)

irr_drain_avg <- irr_drain %>%                         # Aggregate data
  group_by(Date, code, TREATMENT) %>% 
  dplyr::summarize(mlday = mean(mlday),N_level=mean(N_level_p),EC=mean(EC)) %>% 
  as.data.frame()
irr_drain_avg$N_level<-as.factor(irr_drain_avg$N_level)
irr_drain_avg$EC<-as.factor(irr_drain_avg$EC)

  irr<-filter(irr_drain_avg, code=="i")
drain<-filter(irr_drain_avg, code=="d")
irr <- irr[order(irr$Date, irr$TREATMENT), ]
drain <- drain[order(drain$Date, drain$TREATMENT), ]
ET<-irr$Date
ET$TREATMENT<-irr$TREATMENT

drain$ET<-irr$mlday-drain$mlday
as.data.frame(ET)




```

# Plots


```{r echo=FALSE, message=F, fig.cap="Irrigation and drainage", warning=F}
ggplot(data=irr_drain_avg)+
  geom_point(aes(x=Date, y=mlday, color=TREATMENT,shape=code))+
  scale_x_datetime(labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()
```

```{r }



ggplot(data=irr_drain_avg)+
  geom_point(aes(x=Date, y=mlday, color=TREATMENT,shape=code))+
  scale_x_datetime(labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw() 

ggplot(data=drain)+
  geom_point(aes(x=Date, y=mlday, color=TREATMENT))+
    geom_point(aes(x=Date, y=ET, color=TREATMENT),shape=6)+
  scale_x_datetime(labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
geom_point(data=irr,aes(x=Date, y=mlday),color="black")


ggplot(data=drain)+
 geom_point(aes(x=Date, y=mlday, color=N_level, shape=EC))+
  geom_point(data=irr,aes(x=Date, y=mlday),color="black", shape = 3)+
  scale_x_datetime(labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
  labs(y= "Irrigation and Drainage (ml/day)",color="N level (%)",shape="EC (dS/m)")
ggsave("Irr_Drain.png")
```

```{r }
drain_irrET<-full_join(drain,irr,by=c("Date","TREATMENT","EC","N_level"))

ggplot(data=drain_irrET)+
geom_point(aes(x=Date, y=ET, color=N_level, shape=EC))+
geom_point(aes(x=Date, y=mlday.y,fill = "Irrigation"),shape = 1)+
scale_x_datetime(labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
#scale_y_continuous(limits = c(0, 700))
  #geom_point(data=irr,aes(x=Date, y=mlday),color="black", shape = 1)+
  labs(y= "Irrigation and Evapotranspiration (ml/day)", fill = "Irrigation", shape="ET by EC (dS/m)",color="ET by N level (%)")
ggsave("ET_irr.png")

```

```{r }

# this is the plot of ET as a function of N level as bars for each EC treatment
# we can make it also with points and make a non-linear model
# we should also make it with the total seasonal ET and not only with one date. I used only
# one date, (the last one right now) because it was the first one to show differences.

filtered_d <- drain %>%
  filter(Date %in% c("2024-01-21"))

ggplot(data=filtered_d,aes( y=ET, x=N_level, fill=EC))+
  geom_col(position = "dodge", width = 0.7) +
  labs(fill = "EC") +
  theme_minimal()
  geom_col()
 ggsave("ET_bar.png")

 
 # filtered_d_i <-irr_drain%>%
 # filter(code %in% "d")
 # filtered_d_i <- filtered_d_i %>%
 #  filter(Date %in% c("2024-01-14"))
 # 
 # ggplot(data=filtered_d_i,aes( y=mlday, x=TREATMENT))+
 #  geom_col(position = "dodge", width = 0.7) +
 #    stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.2)+
 #   theme_minimal()
 # ggsave("drain_bar.png")

```


# plots for the irrigation part
```{r }
irrFilter<- raw_data_irr %>%
  filter(targetEC %in% c("0.9", "4"))
ggplot(data=irrFilter)+
 geom_point(aes(x=Date, y=Nmgl_irr, color=N_level_p))+
   geom_line(aes(x=Date, y=Nmgl_irr, color=N_level_p, linetype=targetEC.n ))+
  #geom_point(data=irr,aes(x=Date, y=mlday),color="black", shape = 3)+
  scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
  labs(y= "Irrigation N Concentration (mg/l)",color="N level (%)", linetype="EC (dS/m)")
ggsave("Irr_N.png")

ggplot(data=raw_data_irr)+
 geom_point(aes(x=targetN_irr, y=Nmgl_irr))
ggsave("targetN.png")

ggplot(data=raw_data_irr)+
 geom_point(aes(x=targetEC, y=Ecirr))
ggsave("targetEC.png")
```
# plots for the drainage quality
```{r }
# irrFilter<- raw_data_irr %>%
#   filter(targetEC %in% c("0.9", "4"))
drainEC$N_level_p<-as.factor(drainEC$N_level_p)
drainEC$EC<-as.factor(drainEC$EC)
drainN$N_level_p<-as.factor(drainN$N_level_p)
drainN$EC<-as.factor(drainN$EC)
drain_NEC_avg$EC<-as.factor(drain_NEC_avg$EC)
drain_NEC_avg$N_level_p<-as.factor(drain_NEC_avg$N_level_p)
```

```{r }

ggplot(data=drain_NEC_avg)+
 geom_point(aes(x=Date, y=drainEC, color=N_level_p,shape=EC))+
     scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
  facet_wrap(~N_level_p)+
  labs(y= "Drainage EC (dS/m)",color="N level (%)", shape="EC (dS/m)")
#ggsave("Drain_EC.png")


ggplot(data=drain_NEC)+
 geom_smooth(aes(x=Date, y=drainEC, color=TREATMENT))+
     scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
   labs(y= "Drainage EC (dS/m)",color="N level (%)", shape="EC (dS/m)")


ggplot(data=filter(drain_NEC, drainN_NO3mgl<700))+
 geom_smooth(aes(x=Date, y=drainN_NO3mgl, color=TREATMENT))+
     scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
  scale_y_continuous(limits = c(0, 300))+
  #facet_wrap(~N_level_p)+
  labs(y= "Drainage N-NO3 Concentration (mg/l)",color="N level (%)", shape="EC (dS/m)")
#ggsave("Drain_N.png")


ggplot(data=drain_NEC_avg)+
 geom_point(aes(x=Date, y=drainEC, color=TREATMENT))+
     scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()+
  facet_wrap(~N_level_p)+
  labs(y= "Drainage EC (dS/m)",color="N level (%)", shape="EC (dS/m)")

ggplot(data=drain_NEC_avg)+
 geom_point(aes(x=Date, y=drainN_NO3mgl, color=TREATMENT))+
     scale_x_datetime(limits = c(as.POSIXct("2023-09-01"), as.POSIXct("2023-12-31")),labels = date_format("%d-%m-%y"),breaks = date_breaks("1 months"))+  theme_bw()
```

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#-----------------------------

loadfonts(device = "win")

data_1 <- read_excel("eoso.xlsx",sheet = "07022024")

labels <- read_excel("eoso.xlsx",sheet = "labels")


data_long <- data_1 |>
  pivot_longer(
    6:8,
    names_to = "Variable",
    values_to = "Value"
  )
variables <- unique(data_long$Variable)

stats_vars <- data_long |>
  group_by( treatment, Variable) |>
  summarise(
    Mean = mean(Value, na.rm = T),
    Sd = sd(Value, na.rm = T),
    N = n(),
    tcalc = qt(0.975, N-1),
    eem = tcalc*Sd/sqrt(N),
    Min = Mean - eem,
    Max = Mean + eem
  )


groups <- data.frame(
  treatment = 0,
  Variable = "x",
  groups = "x"
)

  for(j in seq_along(variables)){
    df <- data_long |>
      filter(Variable == variables[j])
    mod1 <- aov(Value ~ treatment, df)
    grps_mod1 <- HSD.test(mod1, "treatment")$groups
    grps_mod1 <- grps_mod1 |>
      rownames_to_column(var = "treatment") |>
      select(-2) |>
      mutate(
        Variable = variables[j]
      ) 
    groups <- rbind(groups, grps_mod1)
  }

groups <- groups |>
  filter(treatment != 0)

stats_vars$treatment <- as.character(stats_vars$treatment)
stats_vars <- left_join(stats_vars, groups)

for(i in seq_along(variables)){
  df <- data_long |>
    filter(Variable == variables[i])
  stats_vars_filtrados <- stats_vars |>
    filter(Variable == variables[i])
    stats_vars_filtrados$treatmentC <- as.numeric(stats_vars_filtrados$treatment) 
  name1 <- paste0("graphics/all/",variables[i],"_box.png")
  name3 <- paste0("graphics/all/",variables[i],"_bars.png")
  label <- paste0(labels[i,2])
  graphic1 <- ggplot(df, aes(x = factor(treatment), y = Value)) +
    geom_boxplot() +
    theme_apa() +
    labs(x = "Treatment", y = label) +
    theme(
      legend.position = "bottom"
    )
  grapchic3 <- ggplot(stats_vars_filtrados, aes(x = treatmentC, Mean)) +
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=Min, ymax=Max),
                  width=.3,
                  position=position_dodge(.9)) +
#scale_fill_grey(start = 0.3, end = .8)  +
    theme_apa() +
    ylab(label) +
    xlab("Treatment") +
    theme(
      legend.position = "bottom"
    )
  maximum = 1.1*(max(stats_vars_filtrados$Mean) + max(stats_vars_filtrados$eem))
  graphic3 <- grapchic3 + geom_text(data = stats_vars_filtrados, aes(x = treatment, y = maximum, label = groups),
                                   position = position_dodge(width = .75))
  ggsave(name1, graphic1, width = 8, height = 4)
  ggsave(name3, graphic3, width = 8, height = 4)
}
writexl::write_xlsx(x = stats_vars, "stats_all.xlsx")





